---
title: "ESM 244 - Lab 7 - Part 2 - Spatial Kriging"
author: "Karina Johnston"
date: "2/17/2022"
output: html_document
---

```{r setup, include = TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

# Attach packages
library(tidyverse)
library(here)

# For mapping
library(sf) # simple feature mapping
library(tmap) # interactive map
library(gstat) # geospatial statistics
library(stars) # geospatial raster package

# originally had separate vector and raster packages - used to be clunky, vector turned into sf package, stars turned into the new raster package - spatio-temporal arrays?

```

## Lab 7 - Part 2 - Spatial Kriging

**Data source:** California Department of Water Resources. (2016). Groundwater Depth, California, Fall 2013. California Department of Water Resources. Available at: http://purl.stanford.edu/py212tb3061.

These data are measurements for depth to groundwater recorded in 2013. We'll read in the layer with `read_sf()`, check the CRS, then make a quick plot to start

Read in the California depth to groundwater data


```{r}

# Read in the CA county data (TIGER shapefile):
ca_counties_sf <- read_sf(here("data/ca_counties"), 
                          layer = "CA_Counties_TIGER2016") %>% 
  janitor::clean_names() %>% # clean up the headers
  select(name)

# Check the projection
st_crs(ca_counties_sf)


# read in the groundwater depth data
# .gpkg = geopackage file - more condensed way of storing and only one file!

ca_dgw_sf <- read_sf(here("data/ca_dgw", 
                          "F2013_DBGS_Points_20150720_093252.gpkg")) %>% 
  janitor::clean_names()

# check CRS
st_crs(ca_dgw_sf)

ca_dgw_sf <- st_transform(ca_dgw_sf, st_crs(ca_counties_sf))

```

map it to check out the raw data

```{r}
ggplot() +
  geom_sf(data = ca_counties_sf) +
  geom_sf(data = ca_dgw_sf, aes(color = dgbs)) + # map color to groundwater depth
  theme_classic()

```

interactive map using `tmap`

```{r}
tmap_mode(mode = "view")

tm_shape(ca_dgw_sf) +
  tm_dots("dgbs", palette = "Blues")
# note - no CA map here, Leaflet does that for us

### check out all the palettes:
# tmaptools::palette_explorer() ### need package shinyjs
```

clip the map to San Joaquin county

```{r}
sj_county_sf <- ca_counties_sf %>% 
  filter(name == "San Joaquin")

# now keep groundwater obs in just that county
# take two spatial data sets and intersect them - drops anything with one of them outside the layer - take only the overlap

sj_depth_sf <- ca_dgw_sf %>% 
  st_intersection(sj_county_sf)

```


```{r}
ggplot() +
  geom_sf(data = sj_county_sf) +
  geom_sf(data = sj_depth_sf, aes(color = dgbs)) +
  theme_classic() +
  scale_color_viridis_c()

```

now use the map from above to create a variogram that uses interpolation to connect the data points spatially

but first, need to double check for overlap in points - it will mess up the variogram!
called: "spatial singularity"

#### check for duplicate locations

```{r}
well_duplicates <- sj_depth_sf %>% # create a new df
  janitor::get_dupes(latitude, longitude) # check for multiples

# now just drop them (could average them or something else, but going easy for now)
sj_depth_sf <- sj_depth_sf %>% 
  filter(!local_well %in% well_duplicates$local_well) # drop the ones from the duplicates df

```

### Make a Variogram

```{r}
# use the gstat package

# new df, variable dgbs as a function of 1, says mean is constant using data - ordinary kriging - know mean is constant but don't know what it is - tell it to do orginary kriging

sj_dgw_vgm <- gstat::variogram(dgbs ~ 1, data = sj_depth_sf) 

plot(sj_dgw_vgm) # not super clean because doesn't level off, but will work with it

# need to know nugget (y-intercept), range (point at which you reach the sill), sill (leveling off - may not be able to see it)
```


```{r}
# create a variogram model by fitting the variogram to the points
# potential models: gaussian, spherical, exponential, others?

sj_dgw_vgm_fit <- gstat::fit.variogram(sj_dgw_vgm,
                                       model = vgm(model = "Gau",
                                                   nugget = 15,
                                                   sill = 3000,
                                                   range = 32000)) # distance in m

# now plot on top of other one

plot(sj_dgw_vgm, sj_dgw_vgm_fit)
# ok fit, but not perfect, would be stronger if we give it values for sill, range, nugget

# now that we've added specifics about the values to the variogram, it's a better fit model

attr(sj_dgw_vgm_fit, "SSErr") # whats the sum of square errors for the particular fit
# can compare the different models
```


### Spatial Kriging

now our model is set up, we can come up with the weighted average for the groundwater depth

```{r}
# set up a grid

# start by setting up a bounding box against the county
sj_grid <- st_bbox(sj_county_sf) %>% 
  st_as_stars(dx = 1000, # distance in x
                     dy = 1000 # distance in y
                     )  %>% 
  st_set_crs(st_crs(sj_county_sf)) %>% 
  st_crop(sj_county_sf)

sj_grid

plot(sj_grid)

```

### Perform Ordinary Kriging

```{r}
sj_dgw_krige <- gstat::krige(formula = dgbs ~ 1, 
                             sj_depth_sf, 
                             sj_grid,
                             model = sj_dgw_vgm_fit)

plot(sj_dgw_krige)
```

convert to a spatial data frame that we can plot

```{r}
krige_df <- as.data.frame(sj_dgw_krige) %>% 
  st_as_sf(coords = c("x", "y")) %>% # convert to simple features object
  drop_na(var1.pred)

```

plot it

```{r}
ggplot(data = krige_df) +
  geom_sf(aes(color = var1.pred)) +
  scale_color_gradient(low = "blue", high = "yellow") +
  theme_classic()

```

### End part 2!

