---
title: "ESM 244 - Lab 7 - Part 2 - Spatial Kriging"
author: "Karina Johnston"
date: "2/17/2022"
output: html_document
---

```{r setup, include = TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

# Attach packages
library(tidyverse)
library(here)

# For mapping
library(sf) # single feature mapping
library(tmap) # interactive map
library(gstat) # geospatial statistics
library(stars) # geospatial raster package

# originally had separate vector and raster packages - used to be clunky, vector turned into sf package, stars turned into the new raster package - spatio-temporal arrays?

```

## Lab 7 - Part 2 - Spatial Kriging

Read in the California depth to groundwater data

**Data source:** California Department of Water Resources. (2016). Groundwater Depth, California, Fall 2013. California Department of Water Resources. Available at: http://purl.stanford.edu/py212tb3061.

These data are measurements for depth to groundwater recorded in 2013. We'll read in the layer with `read_sf()`, check the CRS, then make a quick plot to start

```{r}

# Read in the CA county data (TIGER shapefile):
ca_counties_sf <- read_sf(here("data/ca_counties"), layer = "CA_Counties_TIGER2016") %>% 
  janitor::clean_names() %>% 
  select(name)

# Check the projection
st_crs(ca_counties_sf)



```


