---
title: "ESM 244 - Lab 7 - Spatial Analyses"
author: "Karina Johnston"
date: "2/17/2022"
output: 
  html_document: 
    toc: yes
    theme: cerulean
---

## Lab 7 - Spatial Analyses

```{r setup, include=TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

# Attach packages
library(here)
library(tidyverse)
library(broom)

# Spatial packages
library(sf)
library(tmap) # can turn into an interactive map

```

### Part 1: Spatial Data Wrangling, Visualization

Coordinate reference systems - how to change
focusing on vector data today (not raster, but could do next week)

#### CA County Outlines (polygons)

```{r}
# read in the data

ca_counties_sf <- read_sf(here("data/ca_counties/CA_Counties_TIGER2016.shp")) # note this is a simple features object - not a dataframe 

```


wrangling

```{r}
ca_subset_sf <- ca_counties_sf %>% 
  janitor::clean_names() %>%
  select(county_name = name, land_area = aland)

head(ca_subset_sf) ### WARN AGAINST View()
```

Take a look at `ca_subset_sf`. We should notice something very important about a simple features (sf) object: it just **assumes** you want to keep the spatial information, and you can work with the rest of the data as if it's a non-spatial data frame (and the spatial information just "sticks" - hence the term "sticky geometry"). So even though we only called (and renamed) `name` and `aland` in the `select()` function, we see that the `geometry` column still exists! 

What if we wanted just the dataframe, without the geometry?  Convert to dataframe and select out the geometry column:

```{r}
ca_counties_df <- ca_counties_sf %>%
  as.data.frame() %>%
  select(-geometry)
```


#### Check and set the CRS

Use `st_crs()` to check the existing CRS for spatial data. We see that this CRS is "pseudo-mercator" based on WGS 84 - primarily used for web mapping, not analysis.  WGS84 (epsg:3857), also note proj4 string and WKT definitions.

Need to be in the same Coordinate Reference Systems (CRS)

```{r}
ca_subset_sf %>% st_crs()
ca_subset_sf %>% raster::crs() ### to show proj4 string
```

lets plot it

```{r}
ggplot(data = ca_subset_sf) +
  geom_sf(aes(fill = land_area), color = "black", size = 0.1) +
  theme_void() + # takes everything away except a white background
  scale_fill_gradientn(colors = c("cyan", "blue", "purple"))

```


#### Invasive Red sesbania records (spatial points)

read in the data

```{r}

sesbania_sf <- read_sf(here("data", "red_sesbania", "ds80.shp")) %>% 
  janitor::clean_names()

```


```{r}
# check CRS:

sesbania_sf %>% st_crs()
sesbania_sf %>% raster::crs()

# different coord ref systems

## if you know the EPSG code (4-5 digit #)
sesbania_3857_sf <- st_transform(sesbania_sf, 3857)

## if you don't know EPSG code
sesbania_3857_2_sf <- st_transform(sesbania_sf, st_crs(ca_subset_sf)) # based on crs from other subset data

# check new CRS
sesbania_3857_sf %>% st_crs()
```


#### plot them together


```{r}

ggplot() +
  geom_sf(data = ca_subset_sf) + # this layer will show up first
  geom_sf(data = sesbania_3857_sf, size = 1, color = "red") +
  theme_classic()

```

how many instances of each plant are there in each county?

```{r}
# wrangle to find sesbania observations per county

ca_sesb_sf <- ca_subset_sf %>% 
  st_join(sesbania_3857_sf)

head(ca_sesb_sf)
```

which county has biggest problem with the invasive plant?

```{r}

sesb_counts_sf <- ca_sesb_sf %>% 
  group_by(county_name) %>% 
  summarize(n_records = sum(!is.na(id))) # add column to count when not NA in column, so if no records, will output as zero, and rest will count/sum

head(sesb_counts_sf)
```

#### plot it!

```{r}
ggplot(data = sesb_counts_sf) +
  geom_sf(aes(fill = n_records), color = "black", size = 0.1) +
  scale_fill_gradientn(colors = c("lightgrey", "orange", "red")) +
  theme_minimal() +
  labs(fill = "Number of S. punicea records")

```

#### check out data points just in Solano County

```{r}
## subset the data based on Solano County

solano_sesb_sf <- sesbania_3857_sf %>% 
  filter(county == "Solano")

# only keep the solano polygon
solano_sf <- ca_subset_sf %>% 
  filter(county_name == "Solano")

# plot them together
ggplot() +
  geom_sf(data = solano_sf) +
  geom_sf(data = solano_sesb_sf, color = "darkblue") +
  theme_classic()

```

before we were doing static maps... 

### use the tmap package (interactive)

Sometimes we'll want to make a map interactive so that audience members can zoom in, explore different areas, etc. We can use the {tmap} package to create an interactive map. Let's make one for our California counties (fill aesthetic by land area) with the red sesbania locations on top:


```{r}
# set viewing mode to 'interactive'
tmap_mode(mode = "view")

# start with a static map
tm_shape(ca_subset_sf) + # start with a shape 
  tm_fill("land_area", palette = "BuGn") + # add a fill and other aesthetics
  tm_borders(col = "black") + # change borders
tm_shape(sesbania_3857_sf) +
  tm_dots(size = 0.1)

```

See: 

- [`tmap` vignettes](https://cran.r-project.org/web/packages/tmap/vignettes/tmap-getstarted.html)
- [Chapter 8 in Robin Lovelace's "Geocomputation in R"](https://geocompr.robinlovelace.net/adv-map.html#interactive-maps) 


### end of part 1

Part 2 in other file

